---
title: "NHDPlusHR with hygeo output"
author: "dblodgett@usgs.gov"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nhdplushr_demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=4
)
oldoption <- options(scipen = 9999)
options(scipen = 9999)
options("rgdal_show_exportToProj4_warnings"="none")
```

First, we'll load up three data sets. 
1. the mainstem identifiers from NHDPlusV2 (LevelPathIDs)
2. NHDPlusHR Flowlines which have the NHDPlusV2 LevelPathID as an attribute.
3. Reconciled flowline output from the processing shown in `vignette("sugar_creek_refactor")`

```{r load_data}
library(nhdplusTools)
library(hyRefactor)
library(sf)
library(tidyr)
library(dplyr)
library(hygeo)
src_gpkg <- system.file("gpkg/sugar_creek_fort_mill.gpkg", package = "hygeo")

mr_lp <- st_drop_geometry(read_sf(src_gpkg, "NHDFlowline_Network")) %>%
  select(COMID, LevelPathI)

hr_fline <- read_sf(src_gpkg, "hr_NHDPlusFlowline") %>%
  select(NHDPlusID = COMID, LevelPathI, Hydroseq, mr_LevelPathI)

fline <- read_sf("../tests/testthat/data/sugar_creek_hyRefactor.gpkg", "reconcile")

waterbody_edge_list <- get_waterbody_edge_list(fline,
                                               waterbody_prefix = "fp-")
waterbody <- get_flowpath_data(fline,
                                waterbody_edge_list,
                                catchment_prefix = "cat-")
```

Given these three datasets we need to do a little manipulation to get the identifiers straight. 

First we create a cross walk for mainstem ids. This is a little awkward. See: https://github.com/dblodgett-usgs/hyRefactor/issues/5 Using the cross walk, we can add attributes generate outlet node locations from the NHDPlusHR flowlines and add the mainstem identifier cross walk to them.
```{r}
main_id_xwalk <- select(st_drop_geometry(waterbody),
                        local_id = ID, main_id) %>%
  left_join(get_nhd_crosswalk(fline, catchment_prefix = "cat-"), by = "local_id") %>%
  mutate(COMID = as.integer(COMID)) %>%
  select(-local_id) %>%
  distinct() %>%
  left_join(mr_lp, by = "COMID") %>%
  select(-COMID)

hr_nodes <- st_sf(st_drop_geometry(hr_fline),
                  geom =  st_geometry(nhdplusTools::get_node(hr_fline))) %>%
  left_join(main_id_xwalk, by = c("mr_LevelPathI" = "LevelPathI"))
```

Now we just need to make sure things are in the project we want to use for analysis, the columns of the input points are right, and pass them in.
```{r}
hyl <- st_transform(select(hr_nodes, NHDPlusID, main_id), 5070)

waterbody <- st_transform(waterbody, 5070)

(hyl_out <- get_hydrologic_locaton(hyl, waterbody))
```

Now we can map up the matches we got.

```{r}
hr_matched <- filter(hr_fline, NHDPlusID %in% hyl_out$NHDPlusID)

mapview::mapview(fline, lwd = 4, color = "blue", layer.name = "NHDPlusV2") + 
  mapview::mapview(hr_matched, lwd = 3, color = "red", layer.name = "NHDPlusHR") +
  mapview::mapview(hyl_out, cex = 2, color = "black", layer.name = ("NHDPlusHR Outlets"))

```

```{r teardown, include=FALSE, eval=TRUE}
options(oldoption)
```
